Plant-pollintor networks worldwide are composed by the same specific building

### 2021-08-17 Draft/Draft.Rmd
some edits on draft

Here, we used 60 plant-pollinator networks from 18 different studies and 14 countries, alongside a detailed grouping of plants into reproductive strategies and pollinator species into functional groups. Plants were grouped based on a comprehensive dataset that included floral, reproductive and vegetative traits compiled in Lanuza et al., (unpublished) on a larger set of plant-pollinator networks. Pollinators were grouped into the main taxonomical groups that differed in life form and behaviour. Then, we explored which motifs up to five nodes in these set of plant-pollinator networks were over and under-represented. Finally, we explored if the different plants and pollinators functional groups were over- or under-represented in certain positions.

@@

We have compiled 60 plant-pollinator networks from 18 different studies (**Table S1**). All studies sampled plant-pollinator interactions in natural systems and were selected based on wide geographical coverage and presence of interaction frequency as a measure of interaction strength. In total, there were 503 plant species, 1,111 floral visitors species and 6248 of pairwise interactions registered. For ease of data manipulation plant and floral visitors species names were standardize with the help of the package _taxize_ version 0.9.99 [@chamberlain2020].

@@

First, plant species were grouped thorugh hierarchical clustering into the optimal number of functional groups that summarized the main plant reproductive strategies. We used the trait dataset collated in Lanuza et al., (unpublished) that comprised 1,506 plant species and contained the same 60 plant-pollinator networks used in this study plus some non-weighted and weighted metawebs (see **Table S1** Lanuza et al., unpublished). This dataset consisted on 8 floral, 4 reproductive and 3 vegetative traits excluding traits with high percentage of missing values (over 30%; **Table S2**). We opted to calculate the plant functional groups on this larger set of species because of the higher accuracy when delimiting functional groups with that many variables and species @dolnicar2014. Previous to the clustering analysis, we calculated the distance between the different qualitative and scaled quantitative variables with Gower distance wich can handle mixed data [@gower1971]. For this, we used the function _gowdis_ with method _ward.D2_ from the package _FD_ version 1.0-12 [@laliberte2014]. Finally, we conducted hierarchical clustering with the function _hclust_ from the _R_ stats package version 4.0.5 and calculated the optimal number of clusters with the function _kgs_ from the package _maptree_ version 1.4-7 [@white2009].

Second, floral visitors were grouped into functional groups based on taxonomic rank. We opted to divide floral visitors on the taxonomic rank level and not with functional traits because (i) the main taxonomic orders differ in form and behaviour and (ii) have lower complexity of higher taxonomic ranks (i.e., floral visitors had 6 orders versus plants that had 38). Hence, we grouped floral visitors into 6 functional groups that represented the main life strategies: (i) Hymenoptera-Anthophila (bees), (ii) Hymenoptera-non-Anthophila (other non-bee Hymenoptera), (iii) Syrphidae-Diptera, (iv) non-Syrphidae-Diptera, (v) Lepidoptera and (vi) Coleoptera.

@@

[I may have to think for better headings to guide through analysis and results, specially the second section](Jose)

_Overall motif patterns_

@@

_Motif functional groups position_

@@

[Maybe we need to link this with the plot of the scatterpie](Jose)

@@

**Functional groups**

_Plants_

The hierarchical cluster analysis divided the dataset with 1506 species and 15 traits into five different clusters with different and overlapping characteristics (**Figure S3** and **Figure S4**). The subset of plant species used in this study (N = 503) were distributed evenly across these different larger five groups (see subset of the 503 species labels within cluster for all species in **Figure S3**). The first cluster was dominated by herbs with hermaphrodite flowers with high levels of autonomous selfing and we we refer to this group as "selfing herbs". The second cluster was characterized by small perennial species with a mixed of life forms (trees shrubs and herbs) with outcrossing hermaphroditic flowers and we refer to this group of species as "small outcrossing perennials". The third cluster was dominated by also perennial species with a mixed of life forms too and had large self-incompatible hermaphroditic flowers with high number of ovules and we named this group "self-incompatible perennials with large flowers". The fourth cluster had the tallest species, highest proportion of shrub and tree life forms, dioecious and monoecious breeding system with small flowers but the highest number of flowers per plant and inflorescences, and, we refer to this set of species as "tall plants with small unisexual flowers". Finally, the last cluster of species was dominated by small perennial and shot-lived herbs with long self-compatible zygomorphic flowers unable to self-pollinate and we refer to this group as "short-lived outcrossers with long zygomorphic flowers".

_Floral visitors_

[Describe poll funcional groups?](Jose)

[I'm calling them functional groups because is easier to refer as FG to both but maybe is more accurate to refer them as guilds?](Jose)

>>>>

Here, we used 60 plant-pollinator networks from 18 different studies and 14 countries, alongside functional groups of plants and floral visitors based on functional traits and taxonomic rank, respectively. To obtain plant functional groups, we used a comprehensive dataset that included floral, reproductive and vegetative traits compiled on a larger set of networks (Lanuza et al., unpublished). Floral visitor functional groups were divided by the main taxonomical groups that differed in life form and behaviour. Once we split the different plant-pollinator networks into their motif elements, we explored: (i) over and under-represented motifs; (ii) over and under-represented functional groups in the different motif positions; (iii) the probability of finding plant and floral visitor functional groups in the different motif positions; and, (iv) over- and under- represented functional group combinations within the different motifs.

@@

We have compiled 60 plant-pollinator networks from 18 different studies (**Table S1**). All studies sampled plant-pollinator interactions in natural systems and were selected based on wide geographical coverage and presence of interaction frequency as a measure of interaction strength. In total, there were 503 plant species, 1,111 floral visitors species and 6248 pairwise interactions registered. For ease of data manipulation plant and floral visitors species names were standardize with the help of the package _taxize_ version 0.9.99 [@chamberlain2020].

@@

First, plant species were grouped thorugh hierarchical clustering into the optimal number of functional groups that summarized the main plant reproductive strategies. We used the trait dataset collated in Lanuza et al., (unpublished) that comprised 1,506 plant species and contained the same 60 plant-pollinator networks used in this study plus some non-weighted and weighted metawebs (see **Table S1** Lanuza et al., unpublished). This dataset consisted on 8 floral, 4 reproductive and 3 vegetative traits excluding traits with high percentage of missing values (over 30%; **Table S2**). We opted to calculate the plant functional groups on this larger set of species because of the higher accuracy when delimiting functional groups with that many variables and species @dolnicar2014. Previous to the clustering analysis, we calculated the distance between the different qualitative and scaled quantitative variables with Gower distance wich can handle mixed data [@gower1971]. For this, we used the function _gowdis_ with method _ward.D2_ from the package _FD_ version _1.0-12_ [@laliberte2014]. Finally, we conducted hierarchical clustering with the function _hclust_ from the _R_ stats package version _4.0.5_ and calculated the optimal number of clusters with the function _kgs_ from the package _maptree_ version _1.4-7_ [@white2009].

Second, floral visitors were grouped into functional groups based on taxonomic rank. We opted to divide floral visitors on the taxonomic rank level and not with functional traits because (i) the main taxonomic orders differ in form and behaviour and (ii) have lower complexity of higher taxonomic ranks (i.e., floral visitors had 6 orders versus plants that had 38). Hence, we grouped floral visitors into 6 functional groups, as done similarly in past studies [@fenster2004; @ollerton2009], that represented the main life strategies: (i) Hymenoptera-Anthophila (bees), (ii) Hymenoptera-non-Anthophila (other non-bee Hymenoptera), (iii) Syrphidae-Diptera, (iv) non-Syrphidae-Diptera, (v) Lepidoptera and (vi) Coleoptera.

@@

*Over- and under-represented motifs*

@@

*Over and under-represented functional groups on motif positions*

@@

*Over- and under-represented funtional group combinations of motifs*

@@

**Plant and floral visitor functional groups**

The hierarchical cluster analysis divided the dataset into five different clusters with different and overlapping characteristics (**Figure S3** and **Figure S4**): cluster 1 (selfing herbs) consisted on herbs with hermaphrodite flowers with high levels of autonomous selfing; cluster 2 (small outcrossing perennials) had small perennial species with a mixed of life forms (trees shrubs and herbs) with outcrossing hermaphroditic flowers; cluster 3 (self-incompatible perennials with large flowers) comprised perennial species with a mixed of life forms and had large self-incompatible hermaphroditic flowers with high number of ovules; cluster 4 (tall plants with small unisexual flowers) had the tallest species, highest proportion of shrub and tree life forms, dioecious and monoecious breeding system with small flowers but the highest number of flowers per plant and inflorescences; finally, cluster 5 (short-lived outcrossers with long zygomorphic flowers) was dominated by small perennial and short-lived herbs with long self-compatible zygomorphic flowers that were unable to self-pollinate.

In total, there were 1126 pollinator species and were divided into 9 different taxonomical groups. From highest to lowest number of interactions of the different functional groups with plants: bees (2256), non-syrphid Diptera (1768), syrphids (845), Lepidoptera (437), Coleoptera (432), non-bee Hymenoptera (362), lizards (105), birds (80) and other insects (40).
